Lonely and online: an investigation of social media use and social isolation in China

### 2024-08-27 tandf/Response to reviewers second revision jg edits.qmd

---
title: "Response to reviewers - second revision"
format: pdf
---

## Reviewer 2

*1. Timeliness of Data: The research is based on data collected in 2015, a time when the prevalence of mobile internet and social media platforms might differ significantly from the present landscape. The authors utilized a survey conducted from November 25 to December 2, 2015, as mentioned in the document: "We employed Qualtrics to collect the data...from November 25 to December 2, 2015." Given the rapid evolution of digital communication, the applicability of findings based on a decade-old dataset should be reconsidered.*

We agree that the digital environment changes rapidly, that our data are getting dated, and that we needed to make a case for why we are confident that our theory and results are still relevant today. We make the case that our estimates are actually conservative ones because the new developments in the digital environment should increase the strength of the relationship between social media use and social isolationism. As such, we added to the manuscript a three-pronged defense of the age of the data: 1) Based on our theoretical framework, the development of new social media technologies should only exacerbate the differences between urban and rural variation in the relationship between social media use and feelings of isolation. The new technologies make it even more apparent to rural citizens that they may be missing out - the rise of short video content, improved algorithms that better target users with content and tighten their social networks, better content creation tools, etc. The modern digital environment relative to the one at the time our data were collected should heighten the fear of missing out making our estimates conservative ones. 2) Our questions of digital media consumption are general enough to not be context and time dependent. They are not aimed at specific technologies that are no longer in use. 3) The urbanization of China has increased since the data were collected. This makes citizens living in rural areas increasingly separated from many of their family and friends forced to move to cities for jobs. As such, here too, we expect the divide is only exacerbating the problem. While it is true that more people are moving to cities so they will not have the fear of missing out, those people who are left are less likely to have family and friends nearby.

See footnote 3 for our discussion of this issue.

@@

*2. Mismatch Between Literature Review and Data Collection: The literature review incorporates studies from around 2020, such as the reference to "Valkenburg et al., 2022," yet the data used for analysis is from 2015. This discrepancy may affect the validity of conclusions drawn from contemporary theoretical perspectives when applied to outdated data.*

We are happy to revise our literature review section in ways that the reviewer feels would strengthen the paper but we are a bit unsure of how to proceed given two previous comments by reviewers that we add to the literature review section. While we acknowledge that the previous recommendation and this recommendation are not precisely mutually exclusive, it does suggest that if we remove some references here we may be working against the recommendations of the previous reviewers.

Suggestions as to how to move forward would be welcome here.

> Previous reviewer 1 comment: It is suggested to examine the related theories to answer, why the phenomenon of social media use in Mainland China may be different from others.

> Previous reviewer 2 comment: Nothing in the literature review about social media addiction. I suggest to see: "Social isolation, social support and their relationship with smartphone addiction", "The relationship between addiction to smartphone usage and depression among adults: A crosssectional study", "Smartphone addiction among university students in the light of some variables," and "Examining the effects of motives and gender differences on smartphone addiction."

### 2024-08-27 tandf/Social Isolation in China second revision jg edits.qmd
Second response to reviewers

---
title: "Lonely and online: an investigation of social media use and social isolation in China"
date: '`r format(Sys.time(), "%B %d, %Y")`'
abstract: "Research on the link between social media and well being has lead to ambiguous findings - some research has found a positive, negative, and mixed relationship between the two. We employ the case of China, in which social media is controlled in important ways, as a useful alternative case to gain leverage on these ambiguities. Using data from a nationwide survey of Chinese citizens (n=2292), we principally find that individual, heterogeneous effects are an important driver of the relationship between levels of social media use and reported well-being. Additionally, our results lend support to a FOMO (fear of missing out) mechanism as an important driver of the relationship while not finding support for the mechanism of exposure to negative content on social media as a predictor of well-being. Finally, our findings suggest that institutional context may not play an important role in the social media - social isolation relationship."
keywords: [Social Isolation, Social Media, Digital Communication, China]
format:
  tandf-pdf:
    keep-tex: true 
execute: 
  echo: false
  warning: false
keep-tex: true
bibliography: references.bib
csl: taylor-and-francis-apa.csl
shift-heading-level-by: -1
author:
  - name: Jason Gainous
    affiliations:
      - name: University of Sharjah
        department: Department of Mass Communication
        city: Sharjah
        country: United Arab Emirates
    orcid: 0000-0002-3104-2313
    email: jgainous@sharjah.ac.ae
  - name: Andrew W. MacDonald
    affiliations:
      - name: Duke Kunshan University
        department: Division of Social Sciences
        city: Kunshan
        country: China
    orcid: 0000-0002-4695-1560
    email: andrew.macdonald@dukekunshan.edu.cn
    attributes:
      corresponding: true
---

@@

The data used in this study are original. We designed a survey instrument in English, and translated it into Chinese, to measure a range of concepts including social isolationism and various dimensions of social media and digital information consumption. We employed Qualtrics to collect the data. They randomly selected 2292 respondents from their existing panel[^2] from November 25 to December 2, 2015. This sample size on provides for a roughly $\pm2$ margin of error.[^3] While our sample is selected randomly from Qualtrics' online user panel, it isn't representative of China's entire population due to its internet-based recruitment methods. The sample, however, aligns well with Chinese internet users, crucial for our internet effects research. Compared to national data, our sample skews younger and more educated but is similar in income and gender.[^4] These demographics reflect the global digital divide and are in line with expectations for internet users in developing regions. Our findings are particularly relevant for understanding digital media's impact among active internet users in China.

@@

[^3]: We were able to first collect a small sample ($n=286$) to check the reliability and adjust the instrument after before proceeding with the final data collection. We made three adjustments, none of which are related to the measures used in the current study.

[^4]: See https://data.worldbank.org/country/china (go to the Country Profile), and https://www.cnnic.com.cn/AU/Introduction/Introduction/201208/t20120815_33295.htm.

>>>>

---
title: "Lonely and online: an investigation of social media use and social isolation in China"
date: '`r format(Sys.time(), "%B %d, %Y")`'
abstract: "Research on the link between social media and well being has lead to ambiguous findings - some research has found a positive, negative, and mixed relationship between the two. We employ the case of China, in which social media is controlled in important ways, as a useful alternative case to gain leverage on these ambiguities. Using data from a nationwide survey of Chinese citizens (n=2292), we principally find that individual, heterogeneous effects are an important driver of the relationship between levels of social media use and reported well-being. Additionally, our results lend support to a FOMO (fear of missing out) mechanism as an important driver of the relationship while not finding support for the mechanism of exposure to negative content on social media as a predictor of well-being. Finally, our findings suggest that institutional context may not play an important role in the social media - social isolation relationship."
keywords: [Social Isolation, Social Media, Digital Communication, China]
format:
  tandf-pdf:
    keep-tex: true 
execute: 
  echo: false
  warning: false
keep-tex: true
bibliography: references.bib
csl: taylor-and-francis-apa.csl
shift-heading-level-by: -1
---

@@

The data used in this study are original. We designed a survey instrument in English, and translated it into Chinese, to measure a range of concepts including social isolationism and various dimensions of social media and digital information consumption. We employed Qualtrics to collect the data. They randomly selected 2292 respondents from their existing panel[^2] from November 25 to December 2, 2015.[^3] This sample size on provides for a roughly $\pm2$ margin of error.[^4] While our sample is selected randomly from Qualtrics' online user panel, it isn't representative of China's entire population due to its internet-based recruitment methods. The sample, however, aligns well with Chinese internet users, crucial for our internet effects research. Compared to national data, our sample skews younger and more educated but is similar in income and gender.[^5] These demographics reflect the global digital divide and are in line with expectations for internet users in developing regions. Our findings are particularly relevant for understanding digital media's impact among active internet users in China.

@@

[^3]: While the data are from 2015, we contend that they still contribute to understanding of social isolation and social media use. Based on our theory oulined in this paper, the new developments in the digital environment should, in our view, increase the strength of the relationship found in this paper for three reasons: 1)  The new technologies make it even more apparent to rural citizens that they may be missing out - the rise of short video content, improved algorithms that better target users with content and tighten their social networks, better content creation tools, etc. The modern digital environment relative to the one at the time our data were collected should heighten the fear of missing out making our estimates conservative ones. 2) Our questions of digital media consumption described below are general enough to not be context and time dependent. They are not aimed at specific technologies that are no longer in use. 3) The urbanization of China has increased since the data were collected. This makes citizens living in rural areas increasingly separated from many of their family and friends forced to move to cities for jobs. As such, here too, we expect the divide is only exacerbating the problem. While it is true that more people are moving to cities so they will not have the fear of missing out, those people who are left are less likely to have family and friends nearby.

[^4]: We were able to first collect a small sample ($n=286$) to check the reliability and adjust the instrument after before proceeding with the final data collection. We made three adjustments, none of which are related to the measures used in the current study.

[^5]: See https://data.worldbank.org/country/china (go to the Country Profile), and https://www.cnnic.com.cn/AU/Introduction/Introduction/201208/t20120815_33295.htm.

### 2024-08-27 tandf/Social-Isolation-in-China-second-revision-jg-edits.tex
Second response to reviewers

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Lonely and online: an investigation of social media use and social isolation in China},
  pdfauthor={Jason Gainous; Andrew W. MacDonald},
  pdfkeywords={Social Isolation, Social Media, Digital
Communication, China},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\author{\name{Jason Gainous\textsuperscript{a} and Andrew W.
MacDonald\textsuperscript{b}\thanks{CONTACT Andrew W.
MacDonald. Email: andrew.macdonald@dukekunshan.edu.cn}}
\affil{\textsuperscript{a}Department of Mass Communication, University
of Sharjah, Sharjah, United Arab Emirates;
\textsuperscript{b}Division of Social Sciences, Duke Kunshan
University, Kunshan, China}}

@@

The data used in this study are original. We designed a survey
instrument in English, and translated it into Chinese, to measure a
range of concepts including social isolationism and various dimensions
of social media and digital information consumption. We employed
Qualtrics to collect the data. They randomly selected 2292 respondents
from their existing panel\footnote{Qualtrics recruits a large pool of
  respondents for various survey projects through online advertising.
  Recruits who update their profiles at least once every 6 months are
  randomly invited to participate. These recruits are awarded online
  points that can be exchanged for cash or various other
  country-specific gifts. The number of points is based on the length of
  the survey, and since our survey had over fifty questions, respondents
  received a relatively high number of points.} from November 25 to
December 2, 2015. This sample size on provides for a roughly \(\pm2\)
margin of error.\footnote{We were able to first collect a small sample
  (\(n=286\)) to check the reliability and adjust the instrument after
  before proceeding with the final data collection. We made three
  adjustments, none of which are related to the measures used in the
  current study.} While our sample is selected randomly from Qualtrics'
online user panel, it isn't representative of China's entire population
due to its internet-based recruitment methods. The sample, however,
aligns well with Chinese internet users, crucial for our internet
effects research. Compared to national data, our sample skews younger
and more educated but is similar in income and gender.\footnote{See
  https://data.worldbank.org/country/china (go to the Country Profile),
  and
  https://www.cnnic.com.cn/AU/Introduction/Introduction/201208/t20120815\_33295.htm.}
These demographics reflect the global digital divide and are in line
with expectations for internet users in developing regions. Our findings
are particularly relevant for understanding digital media's impact among
active internet users in China.

@@

\includegraphics{Social-Isolation-in-China-second-revision_files/figure-pdf/fig-social-iso-dist-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-second-revision_files/figure-pdf/fig-var-dist-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-second-revision_files/figure-pdf/fig-sm-use-dist-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-second-revision_files/figure-pdf/fig-urban-dist-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-second-revision_files/figure-pdf/fig-social-isolation-interactions-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-second-revision_files/figure-pdf/fig-resids-human-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-second-revision_files/figure-pdf/fig-resids-online-offline-1.pdf}

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Lonely and online: an investigation of social media use and social isolation in China},
  pdfkeywords={Social Isolation, Social Media, Digital
Communication, China},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\author{\name{\textsuperscript{a} and }
\affil{\textsuperscript{a}
\textsuperscript{b}}}

@@

The data used in this study are original. We designed a survey
instrument in English, and translated it into Chinese, to measure a
range of concepts including social isolationism and various dimensions
of social media and digital information consumption. We employed
Qualtrics to collect the data. They randomly selected 2292 respondents
from their existing panel\footnote{Qualtrics recruits a large pool of
  respondents for various survey projects through online advertising.
  Recruits who update their profiles at least once every 6 months are
  randomly invited to participate. These recruits are awarded online
  points that can be exchanged for cash or various other
  country-specific gifts. The number of points is based on the length of
  the survey, and since our survey had over fifty questions, respondents
  received a relatively high number of points.} from November 25 to
December 2, 2015.\footnote{While the data are from 2015, we contend that
  they still contribute to understanding of social isolation and social
  media use. Based on our theory oulined in this paper, the new
  developments in the digital environment should, in our view, increase
  the strength of the relationship found in this paper for three
  reasons: 1) The new technologies make it even more apparent to rural
  citizens that they may be missing out - the rise of short video
  content, improved algorithms that better target users with content and
  tighten their social networks, better content creation tools, etc. The
  modern digital environment relative to the one at the time our data
  were collected should heighten the fear of missing out making our
  estimates conservative ones. 2) Our questions of digital media
  consumption described below are general enough to not be context and
  time dependent. They are not aimed at specific technologies that are
  no longer in use. 3) The urbanization of China has increased since the
  data were collected. This makes citizens living in rural areas
  increasingly separated from many of their family and friends forced to
  move to cities for jobs. As such, here too, we expect the divide is
  only exacerbating the problem. While it is true that more people are
  moving to cities so they will not have the fear of missing out, those
  people who are left are less likely to have family and friends nearby.}
This sample size on provides for a roughly \(\pm2\) margin of
error.\footnote{We were able to first collect a small sample (\(n=286\))
  to check the reliability and adjust the instrument after before
  proceeding with the final data collection. We made three adjustments,
  none of which are related to the measures used in the current study.}
While our sample is selected randomly from Qualtrics' online user panel,
it isn't representative of China's entire population due to its
internet-based recruitment methods. The sample, however, aligns well
with Chinese internet users, crucial for our internet effects research.
Compared to national data, our sample skews younger and more educated
but is similar in income and gender.\footnote{See
  https://data.worldbank.org/country/china (go to the Country Profile),
  and
  https://www.cnnic.com.cn/AU/Introduction/Introduction/201208/t20120815\_33295.htm.}
These demographics reflect the global digital divide and are in line
with expectations for internet users in developing regions. Our findings
are particularly relevant for understanding digital media's impact among
active internet users in China.

@@

\includegraphics{Social-Isolation-in-China-second-revision-jg-edits_files/figure-pdf/fig-social-iso-dist-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-second-revision-jg-edits_files/figure-pdf/fig-var-dist-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-second-revision-jg-edits_files/figure-pdf/fig-sm-use-dist-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-second-revision-jg-edits_files/figure-pdf/fig-urban-dist-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-second-revision-jg-edits_files/figure-pdf/fig-social-isolation-interactions-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-second-revision-jg-edits_files/figure-pdf/fig-resids-human-1.pdf}

@@

\includegraphics{Social-Isolation-in-China-second-revision-jg-edits_files/figure-pdf/fig-resids-online-offline-1.pdf}
